Distinct snow algae communities along an elevational gradient

### 2019-11-01 manuscript/Casey1.Rmd
wrote the touchy feely field bit, rough draft

@@

>>>>

where we were: there are red and green blooms
where we are: there are multiple types of red and green blooms

@@

### Morphologically diverse snow algae blooms across a variety of habitats

Snow algae was widespread in our study area, beginning on May _ when we first observed snow algae blooming in the trees. Over the course of the season, we collected 309 samples from 33 dates and 13 moutains. We did not observe pink snow above treeline until June 6, despite several visits above treeline before that (site data table). We observed snow algae until the first snows of winter, although by later in the season most of the snow had melted out at lower elevations.

Visually, several patterns in snow algae site characteristics and cell morphology were apparent. Most frequently, we observed pink or red snow, typically in areas recieving high sunlight, dominated by cells that looked similar to *Sanguina nivaloides*, frequently containing green cells of varying shapes and sizes. However, red snow could also be dominated by *Chlainomonas* morphologies, in red snow that was visually indistinguishible from red snow containing primarily *S. nivaloides*. *Chlainomonas* was frequently mixed in with *S. nivaloides* as well (Fig. 1d). We also occasionally observed green snow above treeline, which was dominated by green or yellow oval cells with flanges, similar to those described for *Chloromonas muramotoi*. Colored snow often appeared in snow runnels, frequently at elevations near or below treeline. These runnels containing colored snow formed linear streaks, often with orange on the surface and green snow just below the surface (Fig. 1b). At two sites like this that we visited repeatedly, the runnel began with green snow hidden 1-2 cm below the surface, but on subsequent visits the surface snow had changed color to orange. The subsurface snow contained green, often flagellated algae, while surface snow was dominated by orange cells resembling *Chloro. krienitzii* (Fig. 1e) [@Matsuzaki2015]. Below treeline, orange and green snow was common, often dominated by orange oval cells with short spines, similar to *Chloromonas cf. brevispina* (Fig. 1g). Snow dominated by *Chloromonas cf. brevispina* and *Chloromonas cf. nivalis* was abundnat in shady areas below treeline, often in distinctively bronze colored snow (supp data).

### Diverse *rbc*L reveals species-level snow algae taxonomy

### Snow algae community composition varies with date and elevation

This is supported by our 18S and cell count data. Although we found Chlainomonas

### High elevation, high diversity

### Green snow algae turns orange at the surface
